```{r}
#| label: setup
#| include: false
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  out.width = "100%"
)

library(countdown)

# Load required packages for QR code generation
if (!require("qrcode")) install.packages("qrcode")
library(qrcode)

if (!require("magrittr")) install.packages("magrittr")
library(magrittr)

if (!require("here")) install.packages("here")
library(here)
here::i_am("Presentation/presentation.qmd")

include_local_figure <- function(data_source) {
  knitr::include_graphics(
    path = here::here(
      "Presentation/Materials",
      data_source
    ),
    error = TRUE
  )
}

source(
  here::here("R/set_r_theme.R")
)

# Dynamic project name generation using {here}
project_name <- 
  basename(here::here())

# Create a cleaner display name for the lecture
display_name <- 
  gsub("SSoQE-", "", project_name) %>%
  gsub("_", " ", .) %>%
  tools::toTitleCase()

# Create dynamic URLs
github_url <- 
  paste0(
    "https://github.com/SSoQE/",
     project_name
  )

slides_url <- 
  paste0(
    "https://ssoqe.github.io/", 
    project_name, "/"
  )
```

#  {.title}

<br>

::::: columns
::: {.column width="80%"}
:::

::: {.column width="20%"}
```{r}
#| label: ssoqe-logo-title
include_local_figure("Logos/SSOQE_logo3.png")
```
:::
:::::

::::::: r-hstack
::: {data-id="box1" style="background: `r ssoqe_cols['black']`; width: 135px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: `r ssoqe_cols['cambridge_blue']`; width: 405px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: `r ssoqe_cols['persian_green']`; width: 225px; height: 15px; margin: 0px;"}
:::

::: {data-id="box4" style="background: `r ssoqe_cols['satin_sheen_gold']`; width: 135px; height: 15px; margin: 0px;"}
:::
:::::::

::: {.text-color-white .text-font-heading .text-size-heading2 .text-bold}
`r display_name`
:::

::::: columns
::: {.column width="20%"}
:::

::: {.column .text-right .text-size-body width="80%"}
[👤]()Franka Gaiser

[Science School of Quantitative Ecology 2025]{.text-color-black}

[bit.ly/SSoQE](https://bit.ly/SSoQE){.text-color-black}
:::
:::::

# [Learning Objectives]{.bold .r-fit-text} {.title}

## What we'll cover today

::::: columns
::: {.column .incremental width="45%"}
### 🎯 **Main Topics**

- Introducing DL
- DL in Ecology
- DL Hands-on
- Reflecting on DL
:::

::: {.column .incremental width="55%"}
### ✅ **By the end, you'll be able to:**

- Explain the set up of DL models
- Make decisions on DL model design
- Reflect on the pros and cons of DL models
:::
:::::

# [Introduction to Deep Learning]{.bold .r-fit-text} {.title}

## 🤔⁉️ What is Deep Learning ?

```{r}
#| label: Mentimeter one
include_local_figure("QR/Qrcode_Mentimeter_DL.png")
```

Mentimeter code: **7404 9257**

## The world of AI is manifold

```{r}
#| label: AI-world
include_local_figure("Pictures/01_DL_world_PichlerHartig2022.png")
```

## Facets of Machine Learning

<br> 🧠 **Main idea:** Computer learns from data\
🎯 **Main goal:** Making good predictions for new data\
<br>

::::: columns
::: {.column .incremental .fragment width="50%"}
#### **supervised vs unsupervised learning**

<br>
:::

::: {.column .incremental .fragment width="50%"}
#### **classification vs regression**

<br>
:::
:::::

::::::: columns
::: {.column .incremental .fragment width="25%"}
```{r}
#| label: segmentation
include_local_figure("Pictures/02_Segmentation.png")
```
:::

::: {.column .incremental .fragment width="25%"}
```{r}
#| label: clustering
include_local_figure("Pictures/03_Clustering.png")
```
:::

::: {.column .incremental .fragment width="25%"}
```{r}
#| label: classification
include_local_figure("Pictures/04_Classification.png")
```
:::

::: {.column .incremental .fragment width="25%"}
```{r}
#| label: regression
include_local_figure("Pictures/05_Regression.png")
```
:::
:::::::

## The unique feature of Deep Learning

```{r}
#| label: DL network architecture
include_local_figure("Pictures/06_DLArchitecture.png")
```

## Deep Learning vs Machine Learning

❗Different strengths and use cases

::::: columns
::: {.column .incremental width="50%"}
**Deep Learning**\
[unstructured]{.text-highlight-midnightGreen} data such as images, text, or audio

```{r}
#| label: sattelite landuse
include_local_figure("Pictures/08_sattelitelanduse.png")
```
:::

::: {.column .incremental .fragment width="50%"}
**Machine Learning**\
less data in [structured]{.text-highlight-persianGreen}, tabular manner

```{r}
#| label: table
include_local_figure("Pictures/10_table.png")
```
:::
:::::

## 🤔⁉️ Have you built a Deep Learning model before ?

```{r}
#| label: Mentimeter two
include_local_figure("QR/Qrcode_Mentimeter_DL.png")
```

Mentimeter code: **7404 9257**

# [Examples of Deep Learning in Ecology]{.bold .r-fit-text} {.title}

## 🤔⁉️ Which use cases of deep learning in ecology do you know ?

```{r}
#| label: Mentimeter three
include_local_figure("QR/Qrcode_Mentimeter_DL.png")
```

Mentimeter code: **7404 9257**

## 📸 Object detection and image classification for biodiversity monitoring or habitat mapping

<br>

::::: columns
::: {.column .incremental width="50%"}
```{r}
#| label: camera traps
include_local_figure("Pictures/11_wildlifeAI.png")
```
:::

::: {.column .incremental width="50%"}
```{r}
#| label: habitat mapping
include_local_figure("Pictures/12_habitatmapping.png")
```
:::
:::::

## 🏵️ Multispecies distribution modelling

```{r}
#| label: multispecies
include_local_figure("Pictures/14_Multispecies.png")
```

## 🦖 DeepDive: Estimating global biodiversity patterns through time

```{r}
#| label: deepdive
include_local_figure("Pictures/16_DeepDive.png")
```

## 🧬 Species Delimitation 4.0 

```{r}
#| label: species delimitation 4.0
include_local_figure("Pictures/17_SpeciesDelimitation.png")
```

## 🧠 Knowledge Guided ML

```{r}
#| label: knowledge guided ML
include_local_figure("Pictures/18_KnowledgeGuidedML.png")
```

# [Deep Learning Basics Hands-on]{.bold .r-fit-text} {.title}

# [Deep Learning a Research Frontier]{.bold .r-fit-text} {.title}

## Explainable AI

```{r}
#| label: explainableAI
include_local_figure("Pictures/17_ExplainableAI.png")
```

## Using existing DL models

```{r}
#| label: existingAI
include_local_figure("Pictures/18_ExistingAI.png")
```

## When is it worth it?

- Know-how 🤓
- How much more do we learn?
- Ethics 🤝
- Sustainability 🌱

## 🤔⁉️ How would you rank the mentioned concerns when buidling DL models?

```{r}
#| label: Mentimeter four
include_local_figure("QR/Qrcode_Mentimeter_DL.png")
```

Mentimeter code: **7404 9257**

## About me

[Franka Gaiser]{.bold .r-fit-text .text-color-black} Doctoral Candidate and Research Assistant at [University of Bayreuth](https://www.spowi5.uni-bayreuth.de/en/team/Gaiser/index.php)

- 📧 Franka.Gaiser(at)uni-bayreuth.de
- 🌐 [Franka Gaiser LinkedIn](www.linkedin.com/in/franka-gaiser-a1534a202)
- 🐱 [Github](https://github.com/FGaiser)
- 🆔 [ORCID](https://orcid.org/0009-0003-2600-6603)
